Self Regulation Ontology Retest Data Report

### 2017-11-09 code/SRO_Report.Rmd
cleaning up analyses

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(psych)
library(rmarkdown)
library(psych)
library(stringr)
library(plotly)
library(DT)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
render_this <- function(){rmarkdown::render('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/code/SRO_Report.Rmd', output_dir = '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/reports', html_notebook(toc = T, toc_float = T, code_folding = 'hide'))}
options(scipen = 1, digits = 4)
```

@@

@@

ggsave('Lit_Review_Plot.jpg', plot = p3, device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures/", width = 24, height = 20, units = "in")
rm(p1, p2, p3, p1_legend, mylegend)
```

@@

```{r}
p1_t <- lit_review %>%
  filter(task == 'task') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='#00BFC4')+
  theme_bw()+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'none') + 
  xlab("")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3))+
  scale_size_continuous(range = c(5, 50))+
  geom_vline(xintercept = 0, color = "red", size = 1)

p2_t <- lit_review %>%
  filter(task == 'survey') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='#F8766D')+
  theme_bw()+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'none') + 
  xlab("")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3))+
  scale_size_continuous(range = c(5, 50))+
  geom_vline(xintercept = 0, color = "red", size = 1)

p3_t <- arrangeGrob(p1_t, p2_t, nrow=1)

ggsave('Lit_Review_Plot_t.jpg', plot = p3_t, device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures/", width = 27, height = 48, units = "in", limitsize = FALSE)
rm(p1_t, p2_t, p3_t)
```

@@

```{r echo=FALSE}
#Get variables of interest from Ian's release
tmp1 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables.csv')
tmp2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables_noDDM.csv')
tmp3 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables_EZ.csv')
retest_report_vars = c(names(tmp1), names(tmp2), names(tmp3))
retest_report_vars = unique(retest_report_vars)
lit_rev_vars = unique(lit_review$dv)[which(unique(lit_review$dv) %in% retest_report_vars == FALSE)]
retest_report_vars = c(retest_report_vars, lit_rev_vars)
rm(tmp1, tmp2, tmp3, lit_rev_vars)
```

@@

@@

```{r}
rel_df <- data.frame(spearman = rep(NA, length(numeric_cols)),
                     icc = rep(NA, length(numeric_cols)),
                     pearson = rep(NA, length(numeric_cols)),
                     eta_sq = rep(NA, length(numeric_cols)),
                     sem = rep(NA, length(numeric_cols)),
                     var_subs = rep(NA, length(numeric_cols)),
                     var_ind = rep(NA, length(numeric_cols)),
                     var_resid = rep(NA, length(numeric_cols)))

@@

Average reliability metrics and the number of variables for both the surveys and the tasks
Both Spearman $\rho$'s as well the ICC's are higher for surveys.  
Systematic differences between the two measurements ($\eta^2$) are higher for tasks than surveys (potentially indicating learning effects?).  
Unexpectedly, median standard error of measurement is higher for surveys but the distribution below suggests that this is a side effect of having too many task measures with very low SEMs.
```{r warning=FALSE, message=FALSE}
rel_df %>%
  group_by(task) %>%
  summarise(median_spearman = median(spearman),
            median_icc = median(icc),
            median_pearson = median(pearson),
            median_eta_sq = median(eta_sq),
            median_sem = median(sem),
            num_vars = n()) %>%
  datatable() %>%
  formatRound(columns=c('median_spearman', 'median_icc', 'median_pearson','median_eta_sq', 'median_sem'), digits=3)
```

**Table of reliability point estimates**

```{r}
rel_df %>%
  datatable(filter='top') %>%
  formatRound(columns=c('spearman', 'icc', 'pearson','eta_sq', 'sem'), digits=3)
```

### Histograms of reliability metrics

@@

Task measures have significanlty lower reliability (here checking only on point estimates of ICC).

```{r}
summary(lm(icc~task, rel_df))
```

### Relationship between reliability metrics

@@

With this in mind we reduce the list of task measures to a list of one per task by averaging only the raw measures from all the trials in a task. We chose to reduce the information in this manner to avoid any bias stemming from differential amount of interest and procedures applied to certain tasks over others (e.g. a task can have over 10 measures because it has multiple conditions and we have chosen to fit DDM's for specific conditions while another might only have 2 due to our relative inexperience and lack of interest in it). We check whether the number of trials in a task has a significant effect on these average reliabilities of raw measures as well.

@@

## Bootstrapping

```{r message=FALSE, warning=FALSE}
# boot_df <- read.csv('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/input/bootstrap_merged.csv')

boot_df <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_09-27-2017/bootstrap_merged.csv')

>>>>

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(psych)
library(rmarkdown)
library(psych)
library(stringr)
library(plotly)
library(DT)

render_this <- function(){rmarkdown::render('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/code/SRO_Report.Rmd', output_dir = '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/reports', html_notebook(toc = T, toc_float = T, toc_depth = 2, code_folding = 'hide'))}
options(scipen = 1, digits = 4)
```

```{r, message=FALSE, warning=FALSE}
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

trim <- function (x) gsub("^\\s+|\\s+$", "", x)

@@

## Demographics

```{r}
test_demog <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_09-27-2017/t1_data/demographic_health.csv')

retest_demog <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_09-27-2017/demographic_health.csv')

retest_demog = retest_demog[retest_demog$X %in% test_demog$X,]

names(test_demog)[which(names(test_demog) == 'X')] <-'sub_id'
names(retest_demog)[which(names(retest_demog) == 'X')] <-'sub_id'

summary(retest_demog %>%
          select(Sex, Age))
```

@@

ggsave('Lit_Review_Plot.jpg', plot = p3, device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures/", width = 24, height = 20, units = "in", dpi=100)
rm(p1, p2, p3, p1_legend, mylegend)
```

@@

```{r}
p1_t_legend <- lit_review %>%
  filter(task == 'task') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='black')+
  theme_bw()+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'bottom',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30), 
        legend.text = element_text(size=20), 
        legend.key.width = unit(0.75, "inches"), 
        legend.title = element_text(size=28),
        legend.spacing.x = unit(0.5, "inches")) + 
  guides(size = guide_legend(override.aes = list(size=c(9,18,28))),
         shape = guide_legend(override.aes = list(size=18)))+
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3), name="Type")+
  scale_size_continuous(name = "Sample Size")+
  geom_vline(xintercept = 0, color = "red", size = 1)

p1_t <- lit_review %>%
  filter(task == 'task') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='#00BFC4')+
  theme_bw()+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'none',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30)) + 
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3))+
  scale_size_continuous(range = c(5, 35))+
  geom_vline(xintercept = 0, color = "red", size = 1)

p2_t <- lit_review %>%
  filter(task == 'survey') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='#F8766D')+
  theme_bw()+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'none',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30)) + 
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3))+
  scale_size_continuous(range = c(5, 35))+
  geom_vline(xintercept = 0, color = "red", size = 1)

mylegend<-g_legend(p1_t_legend)

p3_t <- arrangeGrob(arrangeGrob(p1_t, p2_t, nrow=1), mylegend, nrow=2,heights=c(10, 1))

ggsave('Lit_Review_Plot_t.jpg', plot = p3_t, device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures/", width = 27, height = 48, units = "in", limitsize = FALSE, dpi = 72)
rm(p1_t, p2_t, p3_t, mylegend, p1_t_legend)
```

@@

```{r echo=FALSE}
#Get variables of interest from Ian's release
tmp1 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables.csv')
tmp2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables_noDDM.csv')
tmp3 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables_EZ.csv')
retest_report_vars = c(names(tmp1), names(tmp2), names(tmp3))
retest_report_vars = unique(retest_report_vars)
lit_rev_vars = as.character(unique(lit_review$dv)[which(unique(lit_review$dv) %in% retest_report_vars == FALSE)])
retest_report_vars = c(retest_report_vars, lit_rev_vars)
rm(tmp1, tmp2, tmp3, lit_rev_vars)
```

@@

## Demographics reliability

for(i in 1:length(names(test_demog))){
  if(is.numeric(test_demog[,i])){
    numeric_cols <- c(numeric_cols, names(test_demog)[i])
  }
}

demog_rel_df <- data.frame(spearman = rep(NA, length(numeric_cols)),
                     icc = rep(NA, length(numeric_cols)),
                     pearson = rep(NA, length(numeric_cols)))

row.names(demog_rel_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
  demog_rel_df[numeric_cols[i], 'spearman'] <- get_spearman(numeric_cols[i], t1_df = test_demog, t2_df = retest_demog) 
  demog_rel_df[numeric_cols[i], 'icc'] <- get_icc(numeric_cols[i], t1_df = test_demog, t2_df = retest_demog)
  demog_rel_df[numeric_cols[i], 'pearson'] <- get_pearson(numeric_cols[i], t1_df = test_demog, t2_df = retest_demog)
}

demog_rel_df
```

```{r echo=FALSE}
rm(test_demog, retest_demog, demog_rel_df, numeric_cols)
```

@@

rel_df <- data.frame(spearman = rep(NA, length(numeric_cols)),
                     icc = rep(NA, length(numeric_cols)),
                     pearson = rep(NA, length(numeric_cols)),
                     eta_sq = rep(NA, length(numeric_cols)),
                     sem = rep(NA, length(numeric_cols)),
                     var_subs = rep(NA, length(numeric_cols)),
                     var_ind = rep(NA, length(numeric_cols)),
                     var_resid = rep(NA, length(numeric_cols)))

@@

```{r message=FALSE, warning=FALSE}
boot_df <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_09-27-2017/bootstrap_merged.csv')

boot_df = boot_df[boot_df$dv %in% retest_report_vars,]

retest_report_vars[which(retest_report_vars %in% boot_df$dv==FALSE)]

### Histograms of reliability metrics (point estimates)

@@

### Relationship between reliability metrics (point estimates)

@@

We reduce the list of task measures to a list of one per task by averaging only the raw measures from all the trials in a task. We chose to reduce the information in this manner to avoid any bias stemming from differential amount of interest and procedures applied to certain tasks over others (e.g. a task can have over 10 measures because it has multiple conditions and we have chosen to fit DDM's for specific conditions while another might only have 2 due to our relative inexperience and lack of interest in it). We check whether the number of trials in a task has a significant effect on these average reliabilities of raw measures as well.

@@
